Interfering with Noninterference \iftext\thanks{\bcp{Identify applicable funding agencies here (DARPA and NSF and...?)}}\fi

### 2020-02-03 main.tex
Some work before other meetings

The {\em underlying} step function,
$\mathord{\stepstounderfull} \in \MACHS \rightarrow \MACHS$, embodies the
standard RISC-V ISA semantics.  \bcp{It's probably clearer not to put RISCV
  in the notation---for most of the paper, the fact that it's a RISCV really
  doesn't matter.  Moreover, I think we want to make it a partial function,
  for simplicity of definitions in this part, and if we insist on RISCV here
  this will confuse people that know it doesn't have a HALT instruction.}

@@

\leo{We might need to change the intuition behind some of these things.}
A {\em contour} $\contour \in \CONTOURS = \COMPONENTS \rightarrow
\{\HIGH,\LOW\} \times \{\HI, \LO\}$\bcp{We need more mnemonic names for
  these (that help us keep straight when we're talking about confidentiality
  and when integrity), both in the tex and in the way they are displayed.}
is a map from machine components
to confidentiality and integrity levels. $\LOW$ marks locations whose
contents are observable from the outside (by the ``attacker,'' e.g.,
by DMA) and $\HIGH$ marks internal ``hidden'' state that cannot be
observed directly (e.g., processor registers and memory addresses for
temporary variables will usually be hidden). Similarly, $\LO$
locations correspond to low-integrity components (e.g. locations that
can be written to \bcp{directly or indirectly} by the network), while $\HI$ locations should be
protected from such flows.

@@

Given a machine state $\mach$ and a contour $\contour$, a {\em
  variant} of $\mach$ with respect to $\contour$ is any machine state
$\mach'$ whose components agree with $\mach$ on all components marked
$\LOW$ by $\contour$---i.e., if $\contour(\component) = \LOW$, then
$\mach(\component) = \mach'(\component)$.

@@

  For each triple $(\contour, \overline{M}_n, \overline{M'}_{n'})$,
  and for every high-integrity component $\component$ for which
  $\contour(\component) = \HI$, $M_0(\component) = M_n(\component)$
  and $M'_0(\component) = M'_{n'}(\component)$.
\end{itemize}
\leo{In the terminology of the original hyperproperty paper,
  confidentiality is a hyperproperty, integrity is a trace property
  that holds for every trace in a hyperproperty. This is a stronger
  version of integrity than DLM for example}.
%
%\leo{Nested lockstep noninterference:}
%
%Lockstep-Noninterference is a restricted form of traditional
%noninterference, that is nevertheless useful (e.g. to encode taint
%tracking in Section~\leo{Blah}) and easy to test (\leo{More blah}).
%At the same time, lockstepness makes it possible to generalize the
%setting and allow for the contours themselves to change throughout the
%execution.

@@

\leo{What about
  instructions the are halted by the monitor?}  \bcp{Not sure what the
  question is here, but note that an instruction halted by the monitor will
  not appear in the trace.}
\item For forward compatibility we could treat the first unmatched return as
  a halt instruction.
\item The instruction memory is marked $\LOW \times \HI$. That means
  it is observable and can't change.  \bcp{What does this assumption mean,
    precisely?  (I.e., what is ``the instruction memory''?)}
\item The memory space is split in two. One part is marked $\LOW
  \times \LO$: this is the public part (or the one that can be
  accessed by the current process e.g.). The other part is marked
  $\HIGH \times \HI$: it should be completely inaccessible to the
  current process.
\item Separation is noninterference:
  \leo{Exactly the definition above}
\item Depending on what we do with final instructions, the property
  could be parameterized by the ``call/return sites''.
\end{itemize}
\leo{(The $+$ part in the title is a hint that one thing that gets added is that
the PCs of corresponding traces always stay in lock step because of integrity?)}
\bcp{The section titles are not working so well, now that we're being a bit
  less abstract about our definitions.}

@@

\leo{Build up to the interesting thing:}
\begin{itemize}
\item The first unmatched return serves as a halt.  \bcp{What does that mean
formally?}
\item The instruction memory is again marked $\LOW \times \HI$. That
  means it is observable (both traces have the same instructions!) and
  can't change.
\item At the beginning of time\bcp{Not clear what ``beginning of time''
  means when we're talking about hyperproperties---the intuition of time is
  relevant to the traces within the hyperproperty, but not to the
  hyperproperty itself.  Simpler, I think, to view the hyperproperty as a
  starting set of traces (where the whole machine state [registers and PC too,
  right?] is marked $\LOW$), closed under an operator that takes a
  contour+trace+trace and snips out all the subtraces between a call and its
return.  (But: eek---see below.)}, the entire data memory is marked $\LOW
  \times \LO$. That makes for a very unintersting property as all
  variants are necessarily identical.
\item BUT, we close the hyperproperty as follows:
  \begin{itemize}
  \item We need the state description stuff to enforce WBCF and be
    able to talk about ``first unmatched returns''. That is also a
    trace property.
  \item After every call, we keep the part of the (left-hand, wlog
    due to symmetry) trace until its first unmatched return.
  \item We construct a new contour at that point: after every call,
    the local state of the caller becomes inaccessible (marked
    $\HIGH \times \HI$) until a subsequent return.
  \item For every variant of the new initial state with respect to the
    new contour, the contour plus pair of traces is in the
    hyperproperty.
  \end{itemize}
\end{itemize}

@@

Under the assumption that WBCF holds, we can describe LSE as an instance of Lockstep-Noninterference, where the contour at
each step maps the current stack frame to \VIS\ and the remainder of the stack to \HID\.
The observations on which noninterference is defined will (necessarily)  be made on the values in the current stack frame.
\apt{Again, we're conflating \VIS\ vs. \HID\ with \LOW\ vs \HIGH\; should probably be teased apart.}
LSE has two aspects: secrecy (callees cannot read caller variables) and integrity (callees cannot alter caller variables).
Secrecy is enforced by the variant-closure clause of L-N: since caller data is \HID, changing it cannot alter
the callee's observable behavior.  Integrity is enforced why??? \apt{I've lost confidence that it is.}

>>>>

We assume we have an underlying step function that takes machine
states to machine states, $\MACHS \rightharpoonup \MACHS$; we also
assume the only possible partiality is because of a $\HALT$
instruction.

%The
%{\em underlying} step function, $\mathord{\stepstounderfull} \in
%\MACHS \rightharpoonup \MACHS$, embodies the standard RISC-V ISA
%semantics.  \bcp{It's probably clearer not to put RISCV in the
%  notation---for most of the paper, the fact that it's a RISCV really
%  doesn't matter.  Moreover, I think we want to make it a partial
%  function, for simplicity of definitions in this part, and if we
%  insist on RISCV here this will confuse people that know it doesn't
%  have a HALT instruction.}

@@

\leo{We might need to change the intuition behind some of these things.}
A {\em contour} $\contour \in \CONTOURS = \COMPONENTS \rightarrow
\{\HIGHSEC,\LOWSEC\} \times \{\HIGHINT, \LOWINT\}$\bcp{We need more mnemonic names for
  these (that help us keep straight when we're talking about confidentiality
  and when integrity) in the way they are displayed.}
is a map from machine components
to confidentiality and integrity levels. $\LOWSEC$ marks locations whose
contents are observable from the outside (by the ``attacker,'' e.g.,
by DMA) and $\HIGHSEC$ marks internal ``hidden'' state that cannot be
observed directly (e.g., processor registers and memory addresses for
temporary variables will usually be hidden). Similarly, $\LOWINT$
locations correspond to low-integrity components (e.g. locations that
can be written to \bcp{directly or indirectly} by the network), while $\HIGHINT$ locations should be
protected from such flows.

@@

Given a machine state $\mach$ and a contour $\contour$, a {\em
  variant} of $\mach$ with respect to $\contour$ is any machine state
$\mach'$ whose components agree with $\mach$ on all components marked
$\LOWSEC$ by $\contour$---i.e., if $\contour(\component) = \LOWSEC$, then
$\mach(\component) = \mach'(\component)$.

@@

  For each triple $(\contour, \overline{M}_n, \overline{M'}_{n'})$,
  and for every high-integrity component $\component$ for which
  $\contour(\component) = \HIGHINT$, $M_0(\component) = M_n(\component)$
  and $M'_0(\component) = M'_{n'}(\component)$.
\end{itemize}
\leo{In the terminology of the original hyperproperty paper,
  confidentiality is a hyperproperty, integrity is a trace property
  that holds for every trace in a hyperproperty. This is a stronger
  version of integrity than DLM for example}.
%
%\leo{Nested lockstep noninterference:}
%
%Lockstep-Noninterference is a restricted form of traditional
%noninterference, that is nevertheless useful (e.g. to encode taint
%tracking in Section~\leo{Blah}) and easy to test (\leo{More blah}).
%At the same time, lockstepness makes it possible to generalize the
%setting and allow for the contours themselves to change throughout the
%execution.

@@

\leo{What about
  instructions the are halted by the monitor?}  \bcp{Not sure what the
  question is here, but note that an instruction halted by the monitor will
  not appear in the trace.}
\item For forward compatibility we could treat the first unmatched return as
  a halt instruction.
\item The instruction memory is marked $\LOWSEC \times \HIGHINT$. That means
  it is observable and can't change.  \bcp{What does this assumption mean,
    precisely?  (I.e., what is ``the instruction memory''?)}
\item The memory space is split in two. One part is marked $\LOWSEC
  \times \LOWINT$: this is the public part (or the one that can be
  accessed by the current process e.g.). The other part is marked
  $\HIGHSEC \times \HIGHINT$: it should be completely inaccessible to the
  current process.
\item Separation is noninterference:
  \leo{Exactly the definition above}
\item Depending on what we do with final instructions, the property
  could be parameterized by the ``call/return sites''.
\end{itemize}
\leo{(The $+$ part in the title is a hint that one thing that gets added is that
the PCs of corresponding traces always stay in lock step because of integrity?)}
\bcp{The section titles are not working so well, now that we're being a bit
  less abstract about our definitions.}

@@

\leo{Build up to the interesting thing:}
\begin{itemize}
\item The first unmatched return serves as a halt.  \bcp{What does that mean
formally?}
\item The instruction memory is again marked $\LOWSEC \times \HIGHINT$. That
  means it is observable (both traces have the same instructions!) and
  can't change.
\item At the beginning of time\bcp{Not clear what ``beginning of time''
  means when we're talking about hyperproperties---the intuition of time is
  relevant to the traces within the hyperproperty, but not to the
  hyperproperty itself.  Simpler, I think, to view the hyperproperty as a
  starting set of traces (where the whole machine state [registers and PC too,
  right?] is marked $\LOWSEC$), closed under an operator that takes a
  contour+trace+trace and snips out all the subtraces between a call and its
return.  (But: eek---see below.)}, the entire data memory is marked $\LOWSEC
  \times \LOWINT$. That makes for a very unintersting property as all
  variants are necessarily identical.
\item BUT, we close the hyperproperty as follows:
  \begin{itemize}
  \item We need the state description stuff to enforce WBCF and be
    able to talk about ``first unmatched returns''. That is also a
    trace property.
  \item After every call, we keep the part of the (left-hand, wlog
    due to symmetry) trace until its first unmatched return.
  \item We construct a new contour at that point: after every call,
    the local state of the caller becomes inaccessible (marked
    $\HIGHSEC \times \HIGHINT$) until a subsequent return.
  \item For every variant of the new initial state with respect to the
    new contour, the contour plus pair of traces is in the
    hyperproperty.
  \end{itemize}
\end{itemize}

@@

Under the assumption that WBCF holds, we can describe LSE as an instance of Lockstep-Noninterference, where the contour at
each step maps the current stack frame to \VIS\ and the remainder of the stack to \HID\.
The observations on which noninterference is defined will (necessarily)  be made on the values in the current stack frame.
\apt{Again, we're conflating \VIS\ vs. \HID\ with \LOWSEC\ vs \HIGHSEC\; should probably be teased apart.}
LSE has two aspects: secrecy (callees cannot read caller variables) and integrity (callees cannot alter caller variables).
Secrecy is enforced by the variant-closure clause of L-N: since caller data is \HID, changing it cannot alter
the callee's observable behavior.  Integrity is enforced why??? \apt{I've lost confidence that it is.}
